---
title: "HWE_testing"
output: html_notebook
---
#Installing the proper necessary packages
```{r}
#setting the working directory first
setwd("~/Desktop/Ritter_Defense_2019/HWE_removal")

#Using dplyr to manipulate data tables and columns within
if(!require(dplyr)){
    install.packages("dplyr")
    library(dplyr)
}
```


#Preparing the data
```{r}
#loading in all the data

HA_19_HWE <- read.table("HWEtest_HA19_forR", sep="\t", header=TRUE)
HA_9_HWE <- read.table("HWEtest_HA9_forR", sep="\t", header=TRUE)
HA_15_HWE <- read.table("HWEtest_HA15_forR", sep="\t", header=TRUE)
KA_ST5R_HWE <- read.table("HWEtest_KA_ST5R", sep="\t", header=TRUE)
KA_GRBD_HWE <- read.table("HWEtest_KA_GRBD", sep="\t", header=TRUE)
Wurtz_HWE <- read.table("HWEtest_Wurtz", sep="\t", header=TRUE)
```

#Making and comparing loci lists
In linux- I printed out ALL the sites from my original beagle file made.
I will now compare them to the loci list generated here, combine them, and generate a list of markers that do not contain any HWE variants.

```{r}
all_markers_from_beagle <- read.table("all_markers_in_datalabeledcolumn.txt", sep="\t", header=TRUE)
markers_to_be_excluded_0.05 <- read.table("markers_to_be_excluded_0.05_correctlylabeled", sep="\t", header=TRUE)

marker_lists_combined_0.05 <- rbind(all_markers_from_beagle, markers_to_be_excluded_0.05)

mutating_dataframe_0.05 <- marker_lists_combined_0.05 %>% add_count(Position)

#creating dataset without loci that are in both dataframes. Aiming to create a dataset with only loci that are in the original beagle angsd file and NOT in the HWE variant list. Also being sure to NOT include those loci that are only in the HWE variant list.
counts_of_markers_to_include_0.05 <- mutating_dataframe_0.05 %>%
  filter(n==1, Data ==1)
#comparing this to the set I would have had without making sure data was only coming from the angsd dataset
counts_of_markers_to_compare_0.05 <- mutating_dataframe_0.01 %>%
  filter(n==1)

#this is the list you will want to use for angsd!
angsd_ready_list_0.05 <- counts_of_markers_to_include_0.05 %>% select("Position")
save(angsd_ready_list_0.05, file="angsd_ready_list_0.05", ascii= TRUE)
```

Now making the lists of which loci have P-values less than 0.001, and combining them to look at which appear in more than 2 populations.

```{r}
#only looking for pvalues <0.001
HA_19_HWE_significantvalues_.001 <- HA_19_HWE %>% filter(p.value <0.001)
HA_9_HWE_significantvalues_.001 <- HA_9_HWE %>% filter(p.value <0.001)
HA_15_HWE_significantvalues_.001 <- HA_15_HWE %>% filter(p.value <0.001)
KA_ST5R_HWE_significantvalues_.001 <- KA_ST5R_HWE %>% filter(p.value <0.001)
KA_GRBD_HWE_significantvalues_.001 <- KA_GRBD_HWE %>% filter(p.value <0.001)
Wurtz_HWE_significantvalues_.001 <- Wurtz_HWE %>% filter(p.value <0.001)

#combining into one dataframe
HWE_sigfig_all_bound_.001 <- rbind(HA_19_HWE_significantvalues_.001, HA_9_HWE_significantvalues_.001, HA_15_HWE_significantvalues_.001, KA_ST5R_HWE_significantvalues_.001, KA_GRBD_HWE_significantvalues_.001, Wurtz_HWE_significantvalues_.001)

#counting each loci
counts_of_loci_morethan1_.001 <-HWE_sigfig_all_bound_.001 %>% count(Position) %>% filter(n>1)
counts_of_loci_.001 <- HWE_sigfig_all_bound_.001 %>% count(Position)
#comparing numbers
nrow(counts_of_loci_.001)
nrow(counts_of_loci_morethan1_.001)

just_marker_names_.001 <- counts_of_loci_morethan1_.001 %>% select(Position)
just_marker_names_.001
save(just_marker_names_.001, file="markers_to_be_excluded_.001", ascii=TRUE)
head(just_marker_names_.001)
tail(just_marker_names_.001)

all_markers_from_beagle_labelled <- read.table("all_markers_in_datalabeledcolumn.txt", sep="\t", header=TRUE)
markers_to_be_excluded_.001 <- read.table("markers_to_be_excluded_.001_correctlynamed", sep="\t", header=TRUE)

```


Combining the list of loci generated in angsd and the list generated from HWE variants to then separate out which are only appearing once (i.e. removing variants which are revealed because they show up in both lists)

```{r}
marker_lists_combined_0.001 <- rbind(all_markers_from_beagle_labelled, markers_to_be_excluded_.001)

mutating_dataframe_0.001 <- marker_lists_combined_0.001 %>% add_count(Position)

#creating dataset without loci that are in both dataframes. Aiming to create a dataset with only loci that are in the original beagle angsd file and NOT in the HWE variant list. Also being sure to NOT include those loci that are only in the HWE variant list.
counts_of_markers_to_include_0.001 <- mutating_dataframe_0.001 %>%
  filter(n==1, Data ==1)
#comparing this to the set I would have had without making sure data was only coming from the angsd dataset
counts_of_markers_to_compare_0.001 <- mutating_dataframe_0.001 %>%
  filter(n==1)

###this is the angsd datafile I want to be working with
#counts_of_markers_to_include_0.001
angsd_ready_list_0.001 <- counts_of_markers_to_include_0.001 %>% select("Position")
save(angsd_ready_list_0.001, file="angsd_ready_list_0.001", ascii= TRUE)
write.csv(angsd_ready_list_0.001,'angsd_ready_list_0.001.csv', quote=FALSE, row.names=FALSE)
```
